Data Exploration

### 2019-01-30 src/data_exploration.Rmd
monthly vifs

```{r}
gearby_mth <- with(presence, table(month, gear))
write.csv(gearby_mth, file = "../output/bio/gear_by_mth.csv")
gearby_mth
```

@@

```{r subset monthly presab}
janpresab <- subset(presab, month == "1")
febpresab <- subset(presab, month == "2")
marpresab <- subset(presab, month == "3")
aprpresab <- subset(presab, month == "4")
maypresab <- subset(presab, month == "5")
junpresab <- subset(presab, month == "6")
julpresab <- subset(presab, month == "7")
augpresab <- subset(presab, month == "8")
seppresab <- subset(presab, month == "9")
octpresab <- subset(presab, month == "10")
novpresab <- subset(presab, month == "11")
decpresab <- subset(presab, month == "12")
```

@@

>>>>

```{r}
gearby_mth <- with(presence, table(gear, month))
write.csv(gearby_mth, file = "../output/bio/gear_by_mth.csv")
gearby_mth
```

@@

```{r subset monthly presab}
presab$gear <- as.character(presab$gear)
janpresab <- subset(presab, month == "1")
febpresab <- subset(presab, month == "2")
marpresab <- subset(presab, month == "3")
aprpresab <- subset(presab, month == "4")
maypresab <- subset(presab, month == "5")
junpresab <- subset(presab, month == "6")
julpresab <- subset(presab, month == "7")
augpresab <- subset(presab, month == "8")
seppresab <- subset(presab, month == "9")
octpresab <- subset(presab, month == "10")
novpresab <- subset(presab, month == "11")
decpresab <- subset(presab, month == "12")
```

@@

#with gear - for jan there is none because there is only one category of gear

#without gear
vif_nogear_jan <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface, data = janpresab))
write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
vif_nogear_jan

#with gear
vif_gear_feb <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface, data = febpresab))
write.csv(vif_gear_feb, "../output/bio/monthly_vifs/vif_gear_feb.csv", row.names = TRUE)
vif_gear_feb

vif_gear_feb <- read.csv("../output/bio/monthly_vifs/vif_gear_feb.csv", header = TRUE)
vif_gear_feb$GVIF2Dfsq <- vif_gear_feb$GVIF..1..2.Df..^2
write.csv(vif_gear_feb, "../output/bio/monthly_vifs/vif_gear_feb.csv", row.names = TRUE)
vif_gear_feb

#without gear
vif_nogear_feb <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface, data = febpresab))
write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
vif_nogear_feb

#with gear
vif_gear_mar <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = marpresab))
write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
vif_gear_mar

vif_gear_mar <- read.csv("../output/bio/monthly_vifs/vif_gear_mar.csv", header = TRUE)
vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
vif_gear_mar

#without gear
vif_nogear_mar <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = marpresab))
write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
vif_nogear_mar

#with gear
vif_gear_apr <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
vif_gear_apr

vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
vif_gear_apr

#without gear
vif_nogear_apr <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
vif_nogear_apr

#with gear
vif_gear_may <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = maypresab))
write.csv(vif_gear_may, "../output/bio/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
vif_gear_may

#without gear
vif_nogear_may <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = maypresab))
write.csv(vif_nogear_may, "../output/bio/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
vif_nogear_may

#with gear
vif_gear_jun <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
write.csv(vif_gear_jun, "../output/bio/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
vif_gear_jun

#without gear
vif_nogear_jun <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
vif_nogear_jun

#with gear
vif_gear_jul <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = julpresab))
write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
vif_gear_jul

vif_gear_jul <- read.csv("../output/bio/monthly_vifs/vif_gear_jul.csv", header = TRUE)
vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
vif_gear_jul

#without gear
vif_nogear_jul <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = julpresab))
write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
vif_nogear_jul

#with gear
vif_gear_aug <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = augpresab))
write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
vif_gear_aug

vif_gear_aug <- read.csv("../output/bio/monthly_vifs/vif_gear_aug.csv", header = TRUE)
vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
vif_gear_aug

#without gear
vif_nogear_aug <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = augpresab))
write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
vif_nogear_aug

#with gear
vif_gear_sep <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = seppresab))
write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
vif_gear_sep

vif_gear_sep <- read.csv("../output/bio/monthly_vifs/vif_gear_sep.csv", header = TRUE)
vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
vif_gear_sep

#without gear
vif_nogear_sep <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = seppresab))
write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
vif_nogear_sep

#with gear
vif_gear_oct <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
vif_gear_oct

vif_gear_oct <- read.csv("../output/bio/monthly_vifs/vif_gear_oct.csv", header = TRUE)
vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
vif_gear_oct

#without gear
vif_nogear_oct <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
vif_nogear_oct

#with gear
vif_gear_nov <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = novpresab))
write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
vif_gear_nov

vif_gear_nov <- read.csv("../output/bio/monthly_vifs/vif_gear_nov.csv", header = TRUE)
vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
vif_gear_nov

#without gear
vif_nogear_nov <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = novpresab))
write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
vif_nogear_nov

#with gear
vif_gear_dec <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = decpresab))
write.csv(vif_gear_dec, "../output/bio/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
vif_gear_dec

#without gear
vif_nogear_dec <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = decpresab))
write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
vif_nogear_dec
```
